---
output: pdf_document
fontsize: 11pt
geometry: margin=1in
---

```{r, include = FALSE }
knitr::opts_chunk$set( echo=FALSE, # hide code from document
                       warning=FALSE, # hide warnings 
                       message=FALSE) # hide messages 
library(tidyverse)
```

# STAA57 W21 Draft Report 

Group ? (member names)    

Link to RStudio Cloud shared project: https://rstudio.cloud/

***

## Introduction 

(*Description of questions that are being investigated*)



### Data 

(*Description of data used to answer the questions*)


## Analysis 


(*Main body of analysis, containing data summaries, visualizations, hypothesis tests, or modeling*)


### Some Sub-Section Title


(*Graphs should be created in R, not imported from file*)


```{r,  fig.width = 4, fig.height = 4, fig.cap= "Graph Caption" } 
tibble( t = seq(0, 2*pi, by=0.1) ) %>% 
  mutate( x = 16*sin(t)^3, 
          y = 13*cos(t)-5*cos(2*t)-2*cos(3*t)-cos(4*t) ) %>% 
  ggplot( aes( x = x, y = y) ) + 
  geom_polygon( fill = "pink" ) +
  # annotate axes 
  xlab( "abscissa" ) + 
  ylab( "ordinate" ) + 
  ggtitle( "Graph Title (optional)"  ) +
  coord_fixed() 
```

### Some Sub-section Title     


(*R output tables can be formatted nicely in .html with* `knitr::kable()`) 

```{r, echo = FALSE}
tibble( X = c("this", "my"), 
        Y = c("is", "table") ) %>% 
  knitr::kable( caption = "Table Caption")
```



## Summary 

(*Recap of your findings and conclusions*)


***  

## Appendix 

(*Everything here does not count towards page limit*)


