---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
data = read_csv("data/clean_data_sta.csv")
```



```{r}
data = read_csv("data/clean_data_sta.csv")
 time_stats = data  %>% select(Instructor_ID,Student_ID,Duration,Session_ID
                                                       ,Training_Type) %>%
  mutate(Solo_Time = Duration * as.integer(str_detect(Training_Type,"solo")) ,Dual_Time = Duration * as.integer(str_detect(Training_Type,"dual||Ins")),
         CC_Time_Dual = Duration * as.integer(str_detect(Training_Type,"CC_dual")),
         CC_Time_Solo = Duration * as.integer(str_detect(Training_Type,"CC_solo")),
         Ins_Dual_Time = Duration * as.integer(str_detect(Training_Type,"Ins")))%>% group_by(Instructor_ID,Student_ID,Session_ID) %>% 
  summarise(
    Total_Duration = sum(Duration),
    Total_Solo = sum(Solo_Time), Total_Dual = sum(Dual_Time), Total_CC_Dual = sum(CC_Time_Dual),
    Total_CC_Solo = sum(CC_Time_Solo), Total_Ins_Dual = sum(Ins_Dual_Time)
  ) %>%
  summarise(Total_Duration = sum(Total_Duration),Total_Solo = sum(Total_Solo), Total_Dual = sum(Total_Dual), Total_CC_Dual = sum(Total_CC_Dual),
    Total_CC_Solo = sum(Total_CC_Solo), Total_Ins_Dual = sum(Total_Ins_Dual)) %>% ungroup() %>% arrange(desc(Total_Duration))

#time_stats

percents =time_stats %>% mutate(Total_Duration = ifelse(Total_Duration>45,1,Total_Duration/45)*100,Total_Solo = 
                        ifelse(Total_Solo<12,Total_Solo/12,1)*100,Total_Dual = ifelse(Total_Dual<17,Total_Dual/17,1)*100,Total_CC_Dual = ifelse(Total_CC_Dual<3,Total_CC_Dual/3,1)*100 ,Total_CC_Solo = ifelse(Total_CC_Solo<5,Total_CC_Solo/5,1)*100 ,Total_Ins_Dual = ifelse(Total_Ins_Dual<5,Total_Ins_Dual/5,1)*100,overall = Total_Duration*45/87+Total_Solo*12/87+Total_Dual*17/87+Total_CC_Dual*3/87+Total_CC_Solo*5/87+Total_Ins_Dual*5/87)%>%
 arrange(desc(overall))
percents%>%select(Student_ID,overall,Total_Duration,Total_Solo,Total_Dual,Total_CC_Dual,Total_CC_Solo,Total_Ins_Dual)

```


```{r}
Student_Time = data %>% unite("Date", c(Year, Month, Day), sep = "-") %>% 
  mutate(Date = as.Date(Date)) %>% group_by(Instructor_ID,Student_ID) %>% 
  distinct(Session_ID,.keep_all = T) %>% mutate(Diff = Date - lag(Date)) %>%
  summarise(Time = sum(Diff,na.rm=T)) %>% ungroup()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_Duration))+ geom_col() +geom_smooth()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_CC_Dual))+ geom_col() +geom_smooth()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_CC_Solo))+ geom_col()+geom_smooth()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_Dual))+ geom_col()+geom_smooth()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_Solo))+ geom_col()+geom_smooth()
percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>% 
  ggplot(aes(x=Time,y=overall))+ geom_point()+geom_smooth()+labs(title = "Overall Percentage Completion VS Time in Flight School",x = "Time in School",y="Overall % completed")

```
We have analysed how close each student has come to completing their pilot license, based on each category.
Total_Duration is total number of flight hours(45 needed).
Total_Solo is the total number of hours flown solo(12 needed).
Total_Dual is total number of hours flown dual(17 needed).
Total_CC_Dual is total number of dual hours flown crosscountry(3 needed).
Total_CC_Solo is total number of solo hours flown crosscountry(5 needed).
Total_Ins_Dual is the total number of dual instrument hours(5 needed).
We assume each student has done 10 hours of pre-flight training.
We can see only Student# 88 passes the course with 100% completion, however multiple students have come very close to completion. Here completion is taken as a weighted mean of % completion of each category.
Furthermore,
Plotting the total time a student was in flight school against overall percentage completion of flight license, we see that students that come close to completion ,do so in a varying amount of time. Thus we hypothesize that students finish the course on at their own pace.