---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
data = read_csv("data/clean_data_sta.csv")
```



```{r}
data = read_csv("data/clean_data_sta.csv")
 time_stats = data  %>% select(Instructor_ID,Student_ID,Duration,Session_ID
                                                       ,Training_Type) %>%
  mutate(Solo_Time = Duration * as.integer(str_detect(Training_Type,"solo")) ,Dual_Time = Duration * as.integer(str_detect(Training_Type,"dual||Ins")),
         CC_Time_Dual = Duration * as.integer(str_detect(Training_Type,"CC_dual")),
         CC_Time_Solo = Duration * as.integer(str_detect(Training_Type,"CC_solo")),
         Ins_Dual_Time = Duration * as.integer(str_detect(Training_Type,"Ins")))%>% group_by(Instructor_ID,Student_ID,Session_ID) %>% 
  summarise(
    Total_Duration = sum(Duration),
    Total_Solo = sum(Solo_Time), Total_Dual = sum(Dual_Time), Total_CC_Dual = sum(CC_Time_Dual),
    Total_CC_Solo = sum(CC_Time_Solo), Total_Ins_Dual = sum(Ins_Dual_Time)
  ) %>%
  summarise(Total_Duration = sum(Total_Duration),Total_Solo = sum(Total_Solo), Total_Dual = sum(Total_Dual), Total_CC_Dual = sum(Total_CC_Dual),
    Total_CC_Solo = sum(Total_CC_Solo), Total_Ins_Dual = sum(Total_Ins_Dual)) %>% ungroup() %>% arrange(desc(Total_Duration))

#time_stats

percents =time_stats %>% mutate(Total_Duration = ifelse(Total_Duration>45,1,Total_Duration/45)*100,Total_Solo = 
                        ifelse(Total_Solo<12,Total_Solo/12,1)*100,Total_Dual = ifelse(Total_Dual<17,Total_Dual/17,1)*100,Total_CC_Dual = ifelse(Total_CC_Dual<3,Total_CC_Dual/3,1)*100 ,Total_CC_Solo = ifelse(Total_CC_Solo<5,Total_CC_Solo/5,1)*100 ,Total_Ins_Dual = ifelse(Total_Ins_Dual<5,Total_Ins_Dual/5,1)*100,overall = Total_Duration*45/87+Total_Solo*12/87+Total_Dual*17/87+Total_CC_Dual*3/87+Total_CC_Solo*5/87+Total_Ins_Dual*5/87)%>%
 arrange(desc(overall))
percents%>%select(Student_ID,overall,Total_Duration,Total_Solo,Total_Dual,Total_CC_Dual,Total_CC_Solo,Total_Ins_Dual)

```


```{r}
Student_Time = data %>% unite("Date", c(Year, Month, Day), sep = "-") %>% 
  mutate(Date = as.Date(Date)) %>% group_by(Instructor_ID,Student_ID) %>% 
  distinct(Session_ID,.keep_all = T) %>% mutate(Diff = Date - lag(Date)) %>%
  summarise(Time = sum(Diff,na.rm=T)) %>% ungroup()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_Duration))+ geom_col() +geom_smooth()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_CC_Dual))+ geom_col() +geom_smooth()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_CC_Solo))+ geom_col()+geom_smooth()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_Dual))+ geom_col()+geom_smooth()
#percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>%
#  ggplot(aes(x=Time,y=Total_Solo))+ geom_col()+geom_smooth()
percents %>% inner_join(Student_Time,by = c("Student_ID","Instructor_ID")) %>% 
  ggplot(aes(x=Time,y=overall))+ geom_point()+geom_smooth()+labs(title = "Overall Percentage Completion VS Time in Flight School",x = "Time in School",y="Overall % completed")

```
Summary:
We analysed how many students met the flight license requirements, and compared their progress.
We have analysed how close each student has come to completing their pilot license requirement, based on each category.
The percentage of completion of the total hours requirement(45hrs needed), total solo hours requirement(12hrs needed), the total dual hours requirement(17hrs needed),
the total dual cross-country hours requirement(3hrs needed),
the total solo cross-country hours requirement(5hrs needed), and the total instrument hours requirement(5hrs needed).
Overall is the weighted mean of each category (based on hours), which is used to estimate their overall progress percentage towards their flight license.
There could be a bias, as students that have been in the program should have a much higher overall progress.
Thus, we graph the performance of each student against their enrollment duration in the flight school.

Assumption:
We assume each student has completed 10 hours of pre-flight training.

Conclusion:
We observed that only Student# 88 completed all the requirements for the pilot's license, however multiple students have come very close to completion.
Furthermore, from our graph, we see that students that come close to completion, do so in a varying amount of time. Thus we hypothesize that students finish the course at their own pace.
We can also see a linear trend in the beginning of the graph which becomes somewhat logarithmic later. This is due to the high number of beginner students with low completion percentages, and the students that complete the course in a longer period of time.
